
**Reproducible Research: Peer Assessment 1**
  
======================================================

## Loading and preprocessing the data
```{r readingdata}
activity<-read.csv("activity.csv")
activity$date<-as.Date(activity$date,"%Y-%m-%d")
```

## What is mean total number of steps taken per day?
```{r steps}
library(dplyr)
library(ggplot2)
activity1<- activity %>% group_by(date) %>% summarise(stepsum=sum(steps))
```
Histogram of Steps per day
```{r histogram,fig.hight=4}
g<- ggplot(activity1,aes(stepsum))
g+geom_histogram(fill="blue")+ggtitle("Total no. of Steps per Day")
dev.copy(png, file = "plot1.png")
dev.off()
```
Mean and Median total number of steps taken per day
```{r}
meansteps<- mean(activity1$stepsum,na.rm=TRUE)
mediansteps<- median(activity1$stepsum,na.rm=TRUE)
print(meansteps)
print(mediansteps)
```

## What is the average daily activity pattern?
```{r}
steps_interval<- activity %>% group_by(interval) %>% summarise(stepsum=mean(steps,na.rm=TRUE))
plot(steps_interval$interval,steps_interval$stepsum,xlab="Interval",ylab="Steps",main="Average Daily Activity Pattern",type="l")
dev.copy(png, file = "plot2.png")
dev.off()

```

The maximum number of steps and its corresponding interval.
```{r}
maxsteps<-steps_interval%>%filter(stepsum==max(stepsum))
print(maxsteps)
```


## Imputing missing values
To count total NA values in steps
```{r}
NA_values<-sum(is.na(activity$steps))
print(NA_values)
```
Filling missing data with the mean number of steps across all days with available data for that particular interval.
```{r imputingNA}
actmerge<-merge(activity,steps_interval,by="interval") 
for(i in 1:length(actmerge$steps)){
  if(is.na(actmerge$steps[i])){
    actmerge$steps[i]=actmerge$stepsum[i]
  }
    else{
      actmerge$steps[i]
      }
}
newactivity<-actmerge%>%select(-stepsum)
```
Total number of steps per day.
```{r afterimputing}
newactivity1<- newactivity %>% group_by(date) %>% summarise(newstepsum=sum(steps))
g<- ggplot(newactivity1,aes(newstepsum))
g+geom_histogram(fill="green")+ggtitle("Total no. of Steps per Day after imputing NA")
dev.copy(png, file = "plot3.png")
dev.off()
```

New mean and meadian after imputing NA values
```{r}
newmeansteps<- mean(newactivity1$newstepsum,na.rm=TRUE)
newmediansteps<- median(newactivity1$newstepsum,na.rm=TRUE)
print(newmeansteps)
print(newmediansteps)
```
The mean remains the same as before imputation, while the median value increased slightly.

## Are there differences in activity patterns between weekdays and weekends?
Creating a new factor variable in the dataset with two levels -- "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r day}
actmerge1<-newactivity
for(i in 1:length(actmerge1$date)){
  if(weekdays(as.Date(actmerge1$date[i]))=="Saturday"|weekdays(as.Date(actmerge1$date[i]))=="Sunday"){
    actmerge1$day[i]="weekend"
  }
  else{
    actmerge1$day[i]="weekday"
  }
}
```
Panel plot containing a time series plot of the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days 
```{r plot}
weekdata<-actmerge1 %>% group_by(day,interval)%>% summarise(stepmean=mean(steps))
ggplot(weekdata,aes (interval, stepmean)) + geom_line() +facet_wrap(day~.,nrow=2,ncol=1)+ggtitle("Mean Steps by Interval depending on Day")
dev.copy(png,file ="plot4.png")
dev.off()
```


